Too hard to get: the role of probabilistic expectations and cognitive complexity in destructive multi-dimensional reference points

### 2025-03-25 manuscript-apa.qmd
update manuscript

---
title: "Too hard to get: the role of probabilistic expectations and cognitive complexity in destructive multi-dimensional reference points"
shorttitle: "Destructive multi-dimensional reference points"
author:
  - name: Aspen Han
    corresponding: true
    orcid: 0000-0003-1474-7968
    email: xiangyuhan@uchicago.edu
    affiliations:
      - name: University of Chicago
        department: Department of Economics
        address:  1126 E. 59th Street
        city: Chicago
        region: IL
        postal-code: 60637
author-note:
  disclosures:
    study-registration: null
    data-sharing: null
    related-report: null
    conflict-of-interest: null
    financial-support: null
    authorship-agreements: null
    gratitude: null
abstract: "This paper investigates the effects of conflicting reference points across different dimensions of utility on effort exertion. Reference-dependent preferences have become more prevalent in economic analysis, but models so far have assumed additive separability across different dimensions of utility, which implies that agents respond to reference points in each dimension in isolaton from one another. Challenging this assumption, I hypothesize that agents consider multi-dimensional reference points holistically: agents are less responsive to reference points if they have low probabilistic expectations of being able to concurrently achieve them and/or if they have difficulty reconciling them into a single baseline against which to evaluate outcomes. I
refine the Koszegi-Rabin reference-dependent preference model and apply it to examine effort exertion under targets in different performance dimensions of a task. The original and refined model produce distinct predictions for optimal effort exertion, which I test in a real effort experiment. The experiment finds...which bears implications for..."
keywords: [conflicting multi-dimensional reference points, probabilistic expectations, cognitive complexity]
bibliography: bibliography.bib
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk:
    out.width: "75%"
execute:
  echo: false
format:
  apaquarto-pdf:
    pdf-engine: xelatex
    documentmode: man
    keep-tex: true
    list-of-figures: true
    floatsintext: true
    fig-width: 6
    fig-asp: 0.618
    fig-align: center
    include-in-header:
      text: |
        \usepackage{amsmath}
        \usepackage{amssymb}
        \usepackage{enumitem}
        \setlist[itemize]{topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt}
        \setlist[enumerate]{topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt}
        \usepackage{caption}
        \captionsetup[table]{aboveskip=20pt, belowskip=0pt}
        \usepackage{booktabs}
        \usepackage{dcolumn}
        \usepackage{siunitx}
---

@@

```{r}
#| label: tbl-demographics-sumstats
#| tbl-pos: H
#| tbl-cap: "Summary statistics of covariate data"
#| apa-note: "Significant p-values at 10% level are bolded"

@@

target_regtbl <- texreg(reg_models, stars = c(0.01, 0.05, 0.1), digits = 2, include.ci = FALSE,,
                        omit.coef = "genderprefer not to say|race|!raceother|age|income|econ|mouse",
                        custom.model.names = c("Without controls", "With controls"),
                        custom.header = list("Accuracy" = 1:2),
                        include.deviance = FALSE,
                        booktabs = TRUE,
                        caption = NULL)

@@

# Appendix

## Derivation of first-order conditions for experimental theoretical specification

>>>>

---
title: "Too hard to get: the role of probabilistic expectations and cognitive complexity in destructive multi-dimensional reference points"
shorttitle: "Destructive multi-dimensional reference points"
author:
  - name: Aspen Han
    corresponding: true
    orcid: 0000-0003-1474-7968
    email: xiangyuhan@uchicago.edu
    affiliations:
      - name: University of Chicago
        department: Department of Economics
        address:  1126 E. 59th Street
        city: Chicago
        region: IL
        postal-code: 60637
author-note:
  disclosures:
    study-registration: null
    data-sharing: null
    related-report: null
    conflict-of-interest: null
    financial-support: null
    authorship-agreements: null
    gratitude: null
abstract: "This paper investigates the effects of conflicting reference points across different dimensions of utility on effort exertion. Reference-dependent preferences have become more prevalent in economic analysis, but models so far have assumed additive separability across different dimensions of utility, which implies that agents respond to reference points in each dimension in isolaton from one another. Challenging this assumption, I hypothesize that agents consider multi-dimensional reference points holistically: agents are less responsive to reference points if they have low probabilistic expectations of being able to concurrently achieve them and/or if they have difficulty reconciling them into a single baseline against which to evaluate outcomes. I refine the Koszegi-Rabin reference-dependent preference model and apply it to examine effort exertion under targets in different performance dimensions of a task. The original and refined model produce distinct predictions for optimal effort exertion, which I test in a real effort experiment."
keywords: [conflicting multi-dimensional reference points, probabilistic expectations, cognitive complexity]
bibliography: bibliography.bib
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk:
    out.width: "75%"
execute:
  echo: false
format:
  apaquarto-pdf:
    pdf-engine: xelatex
    documentmode: man
    keep-tex: true
    list-of-figures: true
    floatsintext: true
    fig-width: 6
    fig-asp: 0.618
    fig-align: center
    include-in-header:
      text: |
        \usepackage{amsmath}
        \usepackage{amssymb}
        \usepackage{enumitem}
        \setlist[itemize]{topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt}
        \setlist[enumerate]{topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt}
        \usepackage{caption}
        \captionsetup[table]{aboveskip=20pt, belowskip=0pt}
        \usepackage{booktabs}
        \usepackage{dcolumn}
        \usepackage{siunitx}
---

@@

```{r}
#| label: tbl-demographics-sumstats
#| warning: false
#| tbl-pos: H
#| tbl-cap: "Summary statistics of covariate data"
#| apa-note: "Significant p-values at 10% level are bolded"

@@

target_regtbl <- texreg(reg_models, stars = c(0.01, 0.05, 0.1), digits = 2, include.ci = FALSE,
                        omit.coef = "genderprefer not to say|race|!raceother|age|income|econ|mouse",
                        custom.model.names = c("Without controls", "With controls"),
                        custom.header = list("Accuracy" = 1:2),
                        include.deviance = FALSE,
                        booktabs = TRUE,
                        caption = NULL)

@@
